Evaluating computational methods for modeling off-normal operation of gas centrifuge cascades

### 2018-05-25 paper.tex
all fits in 4 pages

\newcommand{\SN}{S$_N$}
\renewcommand{\vec}[1]{\bm{#1}} %vector is bold italic
\newcommand{\vd}{\bm{\cdot}} % slightly bold vector dot
\newcommand{\grad}{\vec{\nabla}} % gradient
\newcommand{\ud}{\mathop{}\!\mathrm{d}} % upright derivative symbol

@@

Fuel cycle simulations, as with any simulation process, do not produce results
without uncertainties.  Those uncertainties have two main sources: modeling
uncertainty due to the simplifications and approximations built in to the
simulator itself, and data uncertainty from the quantities provided by the
user. This work we will be focused a subset the data uncertainty related to
the user input that defines the performance of the facilities being simulated.

@@

This study seeks to understand which measures of facility operation have the
most impact on the uncertainty of fissile material production across a complete
fuel cycle.  Uncertainty propagation using a Monte Carlo method will be applied
to the Cyclus fuel cycle simulator \cite{cyclus} to estimate the effect of
individual facility uncertainties on the output metrics. The same simulation
will be simulated $N$ times, for each simulation parameters affected with an
uncertainty will be randomly determined, the distribution of each output metrics
over the N iterations of the simulation, will be used to estimate their
respective uncertainty.

@@

Cycamore & Separation & Separation efficiency     \\\cmidrule{2-3} 
         & Storage    & Residence Time            \\
\toprule
CyCLASS  & Reactor    & Cycle Length              \\\cmidrule{3-3}
         &            & Power                     \\\cmidrule{3-3}
         &            & Fuel Enrichment (PWR-UOX) \\

@@

\subsection{The Fuel Cycle Scenario}
For demonstration purposes, a simple commercial fuel cycle transition is used,
inspired from the EG23 fuel cycle
of the Nuclear Fuel Cycle Evaluation and Screening Report\cite{ES}.
Pictured in Figure \ref{fig:cycle}, this is a
transition from a \gls{LWR} fleet loaded with \gls{UOX} fuel to a
\gls{SFR} fleet loaded \gls{MOX} fuel, considering a $1\%/$y growth
of the nuclear generated power (Figure \ref{fig:power}).
\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[scale=0.31]{cycle.png}
  \caption{Simplified representation of the material flow between the different facilities,
  with each operational parameter associated with an
  uncertainty labeled in green.}\label{fig:cycle}
\end{figure}

@@

\begin{figure}[b] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.18]{power.png}
    \caption{Electric power generated over time, with the full production,
        the \gls{LWR} contribution, and the \gls{SFR} contribution labeled in blue, orange, and green, respectively.}
    \label{fig:power}
\end{figure}

@@

In this work, a time step of 1 month has been considered, i.e., each output
metric is reported once per month. Additionally, decay processes have not be
taken into account.

@@

The total power generated in the simulation is reviewed to ensure that none of
the compounded uncertainties result in material shorates substantial enough to
cause long term reactor shutdowns.  For example, since \gls{SFR} fuel is built
from reprocessed used fuel, the availability of that used fuel may impact the
capability to build the required \gls{SFR}-\gls{MOX} fuel.  While there may be studies in
which this would be an interesting result, in the context of this work, such
shutdowns would be well-documented in the operational histories of the
facilities and should not be characterized as an uncertainty.  Figure
\ref{fig:power_full} shows that uncertainties in generated power are indeed
very low.

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{power_full}
    \caption{Electric power generated as a function of time.  The black line
        represents the reference calculation and the blue zone
        represents the 1 $\sigma$ uncertainty distribution.}\label{fig:power_full}
\end{figure}

@@

\subsection{Natural Uranium Consumption dedicated to PWR-UOX fabrication}

Figure \ref{fig:unat_full} represents the cumulative natural uranium consumption
dedicated to PWR-UOX fuel fabrication as a function of time.  The uncertainty on
the cumulative uranium consumption grows as expected with the time and the
loading of the \glspl{LWR}.  The increase stops around year 125y with the
decommissioning of the last \glspl{LWR}.

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{unat_full}
    \caption{Cumulative consumption of natural uranium as a function of time.  The black line
        represents the reference calculation and the blue zone
        represents the 1 $\sigma$ uncertainty distribution.}\label{fig:unat_full}
\end{figure}

As expected, cooling time, separation efficiency and thermal power do not
affect the natural uranium consumption for LWR-UOX fuel production.  The uncertainty on the natural uranium
consumption is dominated by the \gls{LWR} cycle length: shorter the cycle
length is, the more fuel will be loaded, and \emph{vice versa}.

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{unat_uncer}
    \caption{Cumulative natural uranium consumption relative uncertainty, with
    the total uncertainty at 1$\sigma$ in black, and the uncertainty
    contribution of each parameter in other colors.}\label{fig:unatr_uncer}
\end{figure}
The relative uncertainty of the total uranium consumption starts at about
$5.5\%$ a slowly drops and stabilise around $3\%$.

Uncertainty fluctuation of the full relative uncertainty can be observed before
year 50.  Those fluctuation are induced by the fuel loading: all reactors at the
beginning of the simulation have synchronised fuel cycles, reloading in the same time step.
This phenomenon disappears
with the gradual decommissioning of the initial reactors, spread over 50 years.

@@

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{pu_full}
    \caption{Fissile (plutonium) inventory as a function of time.  The black line
        represents the reference calculation and the blue zone
        represents the 1 $\sigma$ uncertainty distribution.}\label{fig:pu_full}
\end{figure}

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{pu_uncer}
    \caption{Fissile (plutonium) inventory relative uncertainty, with
    the total uncertainty at 1$\sigma$ in black, and the uncertainty
    contribution of each parameter in other colors.}\label{fig:pu_uncer}
\end{figure}

At the onset of fissile material production, very low mean inventories lead to
unreliable artefacts in the relative uncertainty.
After that, the relative uncertainty
starts at about $5\%$ and grows until year
125, then decreases to $10\%$ at the end of the simulation time (180y).  
Between year 40 and 70, the main contributor to the fissile inventory
uncertainty is the cycle length.  As with the uranium needs, different cycle length
implies different number of fuel loads, so different amount of resources used.  Nevertheless, around year 65 (see Figure
\ref{fig:used_fuel}), the stock of used fuel available for reprocessing becomes
small enough to start limiting the production of fissile material.
The dominant contribution to the uncertainty then slowly transitions from
the cycle length to the fuel cooling time.

@@

It is also interesting to note that, the uncertainty contribution from the cycle
length between year 40 and 70 comes mainly from fuel batch loading
frequencies.  After 70 years this contribution follows the thermal
power contribution closely, suggesting that fuel discharge burnup (i.e.  fuel
discharge composition) is the mechanism of this contribution.

@@

This preliminary study may suggest that the most important contributions to
natural uranium consumption and fissile inventory come from the parameters that
govern the exchange of material beteren facilities, and not
from the physical parameters that describe those facilities, such as thermal power or fuel enrichment. Further
analysis should allow us to understand this further.

@@

Two main points can be deduced from this work.  First, in a transitional
fuel cycle, the uncertainty contributions to an output metric can vary over time
depending of various factors: deployment schedule artefacts or constrictions in
the material flows.

@@

Cycle length impacts the discharge fuel burnup and  the frequency of
the fuel loads.  When they combine to produce in uncertainty of the
material availability, this could result in variations related to time delays that may
contribute to the final uncertainty in ways that are not relevant to soem analyses.
That is, a simple time lag in a particular output that otherwise follows an identical history
could manifest itself as a large variation at any particular time.
Those kind of time related uncertainties will require a further analysis
to understand and measure accurately their
contributions to the total uncertainty.

>>>>

\newcommand{\SN}{S$_N$}
\renewcommand{\vec}[1]{\bm{#1}} %vector is bold italic
\newcommand{\vd}{\bm{\cdot}} % slightly bold vector dot
\newcommand{\grad}{\vec{\nabla}} % gradient
\newcommand{\ud}{\mathop{}\!\mathrm{d}} % upright derivative symbol
\newcommand{\ie}{\emph{i.e.\ }}

@@

Fuel cycle simulations, as with any simulation process, do not produce results
without uncertainties.  Those uncertainties have two main sources: modeling
uncertainty due to the simplifications and approximations built in to the
simulator itself, and data uncertainty from the quantities provided by the
user. This work focuses on a subset the data uncertainty related to
the user input that defines the performance of the simulated facilities.

@@

This study seeks to understand which characteristics of facility operation have the
most impact on the uncertainty of fissile material production across a complete
fuel cycle.  Uncertainty propagation using a Monte Carlo method will be applied
to the Cyclus fuel cycle simulator \cite{cyclus} to estimate the effect of
individual facility uncertainties on the output metrics. The same simulation
will be performed $N$ times, with some input parameters randomly determined in each case.
The distribution of each output metrics over the N trials of the simulation
will be used to estimate their
respective uncertainty.

@@

Cycamore & Separation & Separation efficiency     \\\cmidrule{2-3} 
         & Storage    & Residence Time            \\
\toprule
CyCLASS  & Reactor    & Cycle Length              \\\cmidrule{3-3}
         &            & Power                     \\\cmidrule{3-3}
         &            & Fuel Enrichment (\gls{LWR}-\gls{UOX} \\

@@

\subsection{The Fuel Cycle Scenario}
For demonstration purposes, a simple commercial fuel cycle transition is used,
inspired from the EG23 fuel cycle
of the Nuclear Fuel Cycle Evaluation and Screening Report\cite{ES}.
Pictured in Figure \ref{fig:cycle}, this is a
transition from a \gls{LWR} fleet loaded with \gls{UOX} fuel to a
\gls{SFR} fleet loaded \gls{MOX} fuel, considering a $1\%/$y growth
of the nuclear generated power (Figure \ref{fig:power}).
\begin{figure}[b] % replace 't' with 'b' to force it to be on the bottom
  \centering
  \includegraphics[scale=0.31]{cycle.png}
  \caption{Simplified representation of the material flow between the different facilities,
  with each operational parameter associated with an
  uncertainty labeled in green.}\label{fig:cycle}
\end{figure}

@@

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.18]{power.png}
    \caption{Electric power generated over time, including contributions from each reactor type.}
    \label{fig:power}
\end{figure}

@@

In this work, a time step of 1 month has been considered, \ie each output
metric is reported once per month. Additionally, decay processes have not be
taken into account.

@@

The total power generated in the simulation is reviewed to ensure that none of
the compounded uncertainties result in material shortages substantial enough to
cause long term facility shutdowns.  For example, since \gls{SFR} fuel is built
from reprocessed used fuel, availability of that used fuel may impact the
capability to build the required \gls{SFR}-\gls{MOX} fuel.  While there may be studies in
which this would be an interesting result, in the context of this work, such
shutdowns would be well-documented in the operational histories of the
facilities and should not be characterized as an uncertainty.  Figure
\ref{fig:power_full} shows that uncertainties in generated power are indeed
very low.

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{power_full}
    \caption{Electric power generated as a function of time, wit its uncertainty.}\label{fig:power_full}
\end{figure}

@@

\subsection{Natural Uranium Consumption Dedicated to \gls{LWR}-\gls{UOX} fabrication}

Figure \ref{fig:unat_full} represents the cumulative natural uranium consumption
dedicated to \gls{LWR}-\gls{UOX} fuel fabrication as a function of time.  The uncertainty on
the cumulative uranium consumption grows as expected with the time and the
loading of the \glspl{LWR}.  The increase stops around year 125y with the
decommissioning of the last \glspl{LWR}.

As expected, cooling time, separation efficiency and thermal power do not
affect the natural uranium consumption for \gls{LWR}-\gls{UOX} fuel production.  The uncertainty on the natural uranium
consumption is dominated by the \gls{LWR} cycle length: shorter the cycle
length is, the more fuel will be loaded, and \emph{vice versa}.

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{unat_full}
    \caption{Cumulative consumption of natural uranium as a function of time, with
      associated uncertainty.}\label{fig:unat_full}
\end{figure}

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{unat_uncer}
    \caption{Relative uncertainty of the cumulative natural uranium consumption, including
      individual contributions from each varied parameter.}\label{fig:unatr_uncer}
\end{figure}
The relative uncertainty of the total uranium consumption starts at about
$5.5\%$ a slowly drops and stabilize around $3\%$.

Uncertainty fluctuation of the full relative uncertainty can be observed before
year 50.  Those fluctuation are induced by the fuel loading: all reactors at the
beginning of the simulation have synchronized fuel cycles, reloading in the same time step.
This phenomenon disappears
with the gradual decommissioning of the initial reactors, spread over 50 years.

@@

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{pu_full}
    \caption{Fissile (plutonium) inventory as a function of time, with
      associated uncertainty.}\label{fig:pu_full}
\end{figure}

\begin{figure}[t] % replace 't' with 'b' to force it to be on the bottom
    \centering
    \includegraphics[scale=0.3]{pu_uncer}
    \caption{Relative uncertainty in the fissile (plutonium) inventory, with
      individual contributions from each varied parameter.}\label{fig:pu_uncer}
\end{figure}

At the onset of fissile material production, very low mean inventories lead to
unreliable artifacts in the relative uncertainty.
After that, the relative uncertainty
starts at about $5\%$ and grows until year
125, then decreases to $10\%$ at the end of the simulation time (180y).  
Between year 40 and 70, the main contributor to the fissile inventory
uncertainty is the cycle length.  As with the uranium needs, different cycle length
implies different number of fuel loads, so different amount of resources used.  Nevertheless, around year 65 (see Figure
\ref{fig:used_fuel}), the stock of used fuel available for reprocessing becomes
small enough to start limiting the production of fissile material.
The dominant contribution to the uncertainty then slowly transitions from
the cycle length to the fuel cooling time.

@@

It is also interesting to note that, the uncertainty contribution from the cycle
length between year 40 and 70 comes mainly from fuel batch loading
frequencies.  After 70 years this contribution follows the thermal
power contribution closely, suggesting that fuel discharge burnup (\ie fuel
discharge composition) is the mechanism of this contribution.

@@

This preliminary study may suggest that the most important contributions to
natural uranium consumption and fissile inventory come from the parameters that
govern the exchange of material between facilities, and not
from the physical parameters that describe those facilities, such as thermal power or fuel enrichment. Further
analysis should allow us to understand this further.

@@

Two main points can be deduced from this work.  First, in a transitional
fuel cycle, uncertainty contributions to output metrics can vary over time
depending of various factors: deployment schedule artifacts or material flow constrictions.

@@

Cycle length impacts the discharge fuel burnup and  the frequency of
the fuel loads.  When they combine to produce in uncertainty of the
material availability, this could result in variations related to time delays that may
contribute to the final uncertainty in ways that are not relevant to some analyses.
That is, a simple time lag in a particular output that otherwise follows an identical history
could manifest itself as a large variation at any particular time.
Those kind of time related uncertainties will require a further analysis
to understand and measure accurately their
contributions to the total uncertainty.
